---
title: 'R50 overview'
output:
  html_document:
    df_print: paged
---


```{r echo=FALSE}
library(Seurat)
library(tidyverse)
library(DESeq2)
library(ComplexHeatmap)
library(broom)
library(muscat)
library(limma)
library(UpSetR)
library(corrr)
library(stringr)
library(magrittr)
library(emmeans)
library(ggsignif)
library(escape)
library(Seurat)
drake::loadd(filt_glia)
```

# Source missing functions
```{r echo=FALSE}
source("R/set_ident.R")
source("~/cp_drake/R/build_pseudobulk.R")
source("R/gen_dds.R")
```

# Plot FGF/R50 seurat object
```{r warning=FALSE}
filt_glia[["trt"]] <- if_else(is.na(filt_glia$trt), true = "R50E", false = filt_glia$trt)
Embeddings(filt_glia, reduction = "umap") %>% 
  as_tibble() %>%
  bind_cols(trt = filt_glia$trt,
            ct = as.character(Idents(filt_glia))) -> df

df %>% 
  group_by(ct) %>%
  summarize(um1 = mean(UMAP_1), um2 = mean(UMAP_2)) -> labs

df[sample(nrow(df)),] %>%
  ggplot() +
  aes(x=UMAP_1, y=UMAP_2) +
  geom_point(aes(color=fct_relevel(trt, "PF","R50E","FGF1")), size=.01) +
  theme_void() +
  theme(legend.position = "top") +
  ggsci::scale_color_jama(name=NULL) +
  guides(color = guide_legend(override.aes = list(size = 3))) -> by_trt

df %>%
  ggplot() +
  aes(x=UMAP_1, y=UMAP_2) +
  geom_point(aes(color=ct), alpha=0.5,  shape=".") +
  annotate("text", x=labs$um1, y = labs$um2, label=labs$ct, size=3, 
           color="black", fontface=2) +
  annotate(x=-13, xend=-13, y= -15, yend= -8, colour="black", lwd=0.75, geom="segment", arrow = arrow(length = unit(0.15, "cm"))) +
  annotate(x=-13, xend=-6, y=-15, yend=-15, colour="black", lwd=0.75, geom="segment", arrow = arrow(length = unit(0.15, "cm"))) +
  theme_void() +
  theme(legend.position = "none") +
  coord_cartesian() +
  ggsci::scale_color_igv(name=NULL) -> by_ct

by_ct + by_trt
```

# Check number of cells per celltype per sample
```{r}
sfgf <- as.character(unique(filt_glia$orig.ident)[grepl("FGF", unique(filt_glia$orig.ident))])
fgf_combos <- as.list.data.frame(data.frame(combn(sfgf, 4)))
spf <- as.character(unique(filt_glia$orig.ident)[grepl("PF", unique(filt_glia$orig.ident))])
sr50 <- as.character(unique(filt_glia$orig.ident)[grepl("R50", unique(filt_glia$orig.ident))])


 glia_sce_combos <- map(fgf_combos, function(x) {
  combos <- c(x, spf, sr50)
  sub_glia <- subset(filt_glia, subset = orig.ident %in% combos)
  sce <- as.SingleCellExperiment(sub_glia)
  sce <- prepSCE(sce, 
    kid = "ident", # subpopulation assignments
    gid = "trt",  # group IDs (ctrl/stim)
    sid = "orig.ident", # sample IDs (ctrl/stim.1234)
    drop = F)
  return(sce)
})

glia.sce <- as.SingleCellExperiment(filt_glia)
glia.sce <- prepSCE(glia.sce, 
    kid = "ident", # subpopulation assignments
    gid = "trt",  # group IDs (ctrl/stim)
    sid = "orig.ident", # sample IDs (ctrl/stim.1234)
    drop = F)
knitr::kable(t(table(glia.sce$cluster_id, glia.sce$sample_id)))  %>%
  kableExtra::kable_styling()
```


# Check for differential abundance of cell-types
```{r}
# calculate cluster-sample cell counts
n_cells <- table(glia.sce$cluster_id, glia.sce$sample_id)

# calculate cluster proportions across samples
freqs <- prop.table(n_cells, margin = 1)
x <- as.numeric(freqs)
names(x) <- rep(rownames(freqs), dim(freqs)[2])

# prep. data.frame for plotting
df <- data.frame(
    frequency = as.numeric(freqs), 
    cluster_id = rep(rownames(freqs), dim(freqs)[2]),
    trt_id = rep(str_remove(colnames(freqs), pattern = ".*_"), each = dim(freqs)[1]),
    sample_id = rep(str_remove(colnames(freqs), pattern = "_.*"), each = dim(freqs)[1]))

# test for sig differences in cell type numbers
mixmod <- lme4::lmer(frequency ~ trt_id*cluster_id + (1|sample_id), data=df) %>%
  emmeans::emmeans(pairwise ~ trt_id|cluster_id) %$%
  contrasts %>%
  tidy()

df %>%
  write_csv("diffabund_byct.csv")

knitr::kable(mixmod) %>%
  kableExtra::kable_styling()
```

# visualize differences in cell numbers
```{r fig.width=3, fig.height=6}
ggplot(df) +
  aes(x=cluster_id, y=frequency, fill=fct_relevel(trt_id, "PF","R50E","FGF1")) +
  geom_boxplot(outlier.shape = NA, width=0.5) +
  geom_point(position = position_dodge(0.9), size=0.25) +
  cowplot::theme_minimal_grid() +
  theme(legend.position = "none") +
  labs(x=NULL, y = "% of Cells") +
  scale_y_continuous(labels = scales::percent) +
  geom_signif(y_position=c(.15, .2, .25), xmin=c(1.7, 2, 1.7), xmax=c(1.9, 2.2, 2.2), angle=270, vjust=-0.35, textsize = 3,
              annotation=c("      *", "       ns", "           ***"), tip_length=.01) +
  ggsci::scale_fill_jama(name=NULL) +
  coord_flip() -> diff_ab

diff_ab
```



# Visualize differences across all cell-types and conditions
## this shows us that biggest conserved difference between conditions is found in COPs
```{r warnings=FALSE}
# aggregate counts by cluster and samples

pb_combos <- map(glia_sce_combos, function(x) {
  aggregateData(x,
    assay = "counts", fun = "sum",
    by = c("cluster_id", "sample_id"))
})

# run MDS
pb_mds <- pbMDS(pb)

# plot MDS
pb_mds + 
  scale_shape_manual(name  = "Treatment", values=c(8,1,3)) +
  scale_color_discrete(name = "Celltype") +
  hrbrthemes::theme_ipsum()
```

# Compare FGF1 with Veh
```{r echo=FALSE}
# specify contrast of interest
de_combos <- map(pb_combos, function(x){ 
  ei <- metadata(glia.sce)$experiment_info
  mm <- model.matrix(~ 0 + ei$group_id)
  dimnames(mm) <- list(ei$sample_id, levels(ei$group_id))
  contrast <- makeContrasts("FGF-PF", levels = mm)
  
  # run DS analysis
  res_fgf <- pbDS(x, design = mm, contrast = contrast)

  # extract DE results
  tbl_fgf <- res_fgf$table[[1]]
  
  return(tbl_fgf)
})

# compare results across groups
res_fp <- map(de_combos, function(x) {
  lapply(x, function(u) {
    u <- dplyr::filter(u, p_adj.loc < 0.05)
    dplyr::arrange(u, p_adj.loc)
  }) %>%
    bind_rows()
})

res_fp %>%
  bind_rows(.id="id") %>%
  group_by(id, cluster_id) %>%
  tally() %>%
  ungroup() %>%
  group_by(cluster_id) %>%
  summarise(mean = median(n)) -> n_med_res_fp

  

res_fp %>%
  bind_rows(.id="id") %>%
  select(id, gene, cluster_id) %>%
  group_by(cluster_id, gene) %>% 
  tally() %>% 
  filter(n>14) %>%
  group_by(cluster_id) %>%
  summarise(n=n()) %>%
  bind_rows(data.frame(cluster_id = "Micro", n = 0))-> n_de_resamp_fp


n_de_fp <- n_de_resamp_fp$n
names(n_de_fp) <- n_de_resamp_fp$cluster_id
n_de_fp
```

```{r}
ei <- metadata(glia.sce)$experiment_info
mm <- model.matrix(~ 0 + ei$group_id)
dimnames(mm) <- list(ei$sample_id, levels(ei$group_id))
contrast <- makeContrasts("R50E-PF", levels = mm)

# run DS analysis
res_rp <- pbDS(pb, design = mm, contrast = contrast)

# extract DE results
tbl_rp <- res_rp$table[[1]]

# filter results on padj and logfc
tbl_fil_rp <- lapply(tbl_rp, function(u) {
  u <- dplyr::filter(u, p_adj.loc < 0.05)
  dplyr::arrange(u, p_adj.loc)
})

# nb. of DS genes & % of total by cluster
n_de_rp <- vapply(tbl_fil_rp, nrow, numeric(1))
p_de_rp <- format(n_de_rp / nrow(glia.sce) * 100, digits = 3)
data.frame("# Differentially Expressed" = n_de_rp, "% DE" = p_de_rp, check.names = FALSE) %>%
  kableExtra::kable() %>% 
  kableExtra::kable_styling()
```

```{r}
ei <- metadata(glia.sce)$experiment_info
mm <- model.matrix(~ 0 + ei$group_id)
dimnames(mm) <- list(ei$sample_id, levels(ei$group_id))
contrast <- makeContrasts("FGF-R50E", levels = mm)

# run DS analysis
res_fr <- pbDS(pb, design = mm, contrast = contrast)

# extract DE results
tbl_fr <- res_fr$table[[1]]

# filter results on padj and logfc
tbl_fil_fr <- lapply(tbl_fr, function(u) {
  u <- dplyr::filter(u, p_adj.loc < 0.05)
  dplyr::arrange(u, p_adj.loc)
})

# nb. of DS genes & % of total by cluster
n_de_fr <- vapply(tbl_fil_fr, nrow, numeric(1))
p_de_fr <- format(n_de_fr / nrow(glia.sce) * 100, digits = 3)
data.frame("# Differentially Expressed" = n_de_fr, "% DE" = p_de_fr, check.names = FALSE) %>%
  kableExtra::kable() %>% 
  kableExtra::kable_styling()
```

```{r fig.width=5, fig.height=15}
rbind((n_de_fp[sort(names(n_de_fp))]), (n_de_fr[sort(names(n_de_fr))]), (n_de_rp[sort(names(n_de_rp))])) %>%
  t() %>% 
  data.frame() %>%
  rownames_to_column("cell") %>% 
  pivot_longer(-cell) %>% 
  mutate(name = case_when(name=="X1" ~ "FGF1vPF",
                          name=="X2" ~ "FGF1vR50E",
                          name =="X3" ~ "R50EvPF")) %>% 
  write_csv("deg_for_prism.csv")
  ggplot() +
  aes(x=name, y=value+1, fill=name) +
  geom_col(position=position_dodge()) +
  coord_flip() +
  facet_wrap(.~cell, scales = "free", ncol=1) +
  ggsci::scale_fill_jama(name=NULL) + 
  labs(y="# DE Genes", x= "Cell Type") +
  theme_bw() -> de_comp
de_comp
```
# Look at how top upregulated FGF1 genes look across all samples and R50
```{r warnings=F, fig.width=5, fig.height=10}
ei <- metadata(glia.sce)$experiment_info
mm <- model.matrix(~ 0 + ei$group_id)
dimnames(mm) <- list(ei$sample_id, levels(ei$group_id))
contrast <- makeContrasts("FGF-PF", levels = mm)

# run DS analysis
res_plot <- pbDS(pb, design = mm, contrast = contrast)

pbht <- pbHeatmap(glia.sce, res_plot, top_n = 10) 
pbht
```

# GSVA analysis
```{r fig.height=8, fig.width=7.5}
# normalize data with deseq2
filt_glia$cellbyid <- paste0(filt_glia$orig.ident, "_", Idents(filt_glia))
vec <- factor(filt_glia$cellbyid)
mm_unfilt <- model.matrix(~ 0 + vec)
mm <- mm_unfilt[, colSums(mm_unfilt) > 10]
colnames(mm) <- levels(vec)[colSums(mm_unfilt) > 10]
mat.sparse <- GetAssayData(filt_glia, slot = "counts", assay = "SCT")
mat.summary.mm <- mat.sparse %*% mm
meta <- filt_glia@meta.data[match(colnames(mat.summary.mm), as.character(filt_glia$cellbyid)),]
dds <- DESeqDataSetFromMatrix(countData = mat.summary.mm,
                                colData = meta,
                                design = ~ 0 + trt)
  
# filter genes based on defined thresholds
keep <- Matrix::rowSums(counts(dds) >= 10) > 30
dds <- dds[keep,]
norm <- vst(dds,blind = F)

# get genesets and run gsva
m_df = msigdbr::msigdbr(species = "Mus musculus")#, category = "C6")
m_df %>% 
  filter(gs_subcat %in% c("CP:REACTOME")) -> filt_df
split(x = filt_df$gene_symbol, f = filt_df$gs_name) -> gs_list
x <- GSVA::gsva(assay(norm), gset.idx.list = gs_list, min.sz=10)

# run anova
x %>% 
  data.frame() %>% 
  rownames_to_column("gene_sets") %>% 
  pivot_longer(-gene_sets) %>%
  mutate(name = str_replace_all(name, "[.]", "-"),
         name = str_remove_all(name, "X")) %>%
  left_join(colData(norm) %>% 
               data.frame %>% 
               rownames_to_column("name") %>%
               dplyr::select(name, trt, orig_labels)) %>%
  # filter(trt!="PF") %>% 
  # mutate(trt = ifelse(trt == 'FGF', 1, 0)) %>%
  group_by(gene_sets, orig_labels) %>%
  group_modify(~lm(value~trt, data=.x) %>% 
                 broom::glance()) -> test
  
# extract all significant mapk/erk terms
test %>%
  mutate(padj = p.adjust(p.value)) %>% 
  filter(grepl("MAPK|erk", gene_sets, ignore.case = T)) %>% 
  arrange(padj) %>% 
  filter(padj<0.05) -> sig_gs

# clean up df
sig_gs %>% 
  dplyr::select(gene_sets, padj, orig_labels) %>% 
  pivot_wider(names_from = orig_labels, values_from = padj) %>%
  replace(is.na(.), 1) %>%
  ungroup() %>% 
  mutate(gene_sets = str_replace_all(string = gene_sets, "_", " ") %>%
           str_remove_all("REACTOME"),
         gene_sets = str_to_sentence(gene_sets) %>%
           str_wrap(30)) %>%
  data.frame() %>% 
  pivot_longer(-gene_sets) -> dfplot

# set order for heatmap
dfplot %>%
  filter(name=="Astro") %>% 
  arrange(-value) %>% 
  pull(gene_sets) -> order

# plot heatmap
dfplot %>% 
  ggplot() +
  aes(x=(name), y=fct_relevel(gene_sets, order), fill=ifelse(-log10(value)>1.3, yes = -log10(value), no=0))  +
  geom_tile(colour="white", size=1) +
  #eom_point(aes(size=-log10(value))) +
  theme_bw() +
  cowplot::theme_half_open() +
  theme(axis.text.x = element_text(angle=45, hjust=1, face="bold"),
        axis.text.y = element_text(angle=15, size = 10),
        legend.position = "right") +
  labs(x=NULL, y=NULL) +
  scale_fill_viridis_c(name="val") -> heat 

heat
```
```{r fig.width=20, fig.height=20}
test %>% 
  mutate(padj = p.adjust(p.value)) %>% 
  filter(padj < 0.05) %>% 
  filter(orig_labels=="Tany") %>% 
  arrange(padj) %>%
  ungroup() %>% 
  dplyr::slice(1:20) %>% 
  pull(gene_sets) -> sig_gs_tany
  
# generate boxplot of top mapk terms
x %>% 
  data.frame() %>% 
  rownames_to_column("gene_sets") %>% 
  pivot_longer(-gene_sets) %>%
  mutate(name = str_replace_all(name, "[.]", "-"),
         name = str_remove_all(name, "X")) %>%
  left_join(colData(norm) %>% 
               data.frame %>% 
               rownames_to_column("name") %>%
               dplyr::select(name, trt, orig_labels)) %>%
  filter(orig_labels=="Tany", 
         gene_sets %in% sig_gs_tany) %>%
  mutate(gene_sets = str_replace_all(string = gene_sets, "_", " ") %>%
           str_remove_all("REACTOME"),
         gene_sets = str_to_title(gene_sets) %>%
           str_wrap(30)) %>%
  ggplot() +
  aes(x=fct_relevel(trt, c("PF","R50E","FGF")), y=value, fill=trt) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = .1) +
  facet_wrap(.~gene_sets, scales="free_y", ncol = 4) +
  theme_bw() +
  theme(strip.background = element_rect(fill="white")) +
  labs(y="Scaled Enrichment", x=NULL) -> boxes
```

```{r, fig.height=8, fig.width=6}
# get top gene sets
sig_gs %>% 
  select(gene_sets, padj, orig_labels) %>% 
  pivot_wider(names_from = orig_labels, values_from = padj) %>%
  replace(is.na(.), 1) %>%
  ungroup() %>% 
  data.frame() %>% 
  pivot_longer(-gene_sets) %>%
  filter(name=="Astro") %>%
  arrange(value) %>%
  top_n(6, -value) %>% 
  pull(gene_sets) -> to_plot

# generate boxplot of top mapk terms
x %>% 
  data.frame() %>% 
  rownames_to_column("gene_sets") %>% 
  pivot_longer(-gene_sets) %>%
  mutate(name = str_replace_all(name, "[.]", "-"),
         name = str_remove_all(name, "X")) %>%
  left_join(colData(norm) %>% 
               data.frame %>% 
               rownames_to_column("name") %>%
               select(name, trt, orig_labels)) %>%
  filter(orig_labels=="Astro", 
         gene_sets %in% to_plot) %>%
  mutate(gene_sets = str_replace_all(string = gene_sets, "_", " ") %>%
           str_remove_all("REACTOME"),
         gene_sets = str_to_title(gene_sets) %>%
           str_wrap(30)) %>%
  ggplot() +
  aes(x=fct_relevel(trt, c("PF","R50E","FGF")), y=value) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = .1) +
  facet_wrap(.~gene_sets, scales="free_y", ncol = 2) +
  theme_bw() +
  theme(strip.background = element_rect(fill="white")) +
  labs(y="Scaled Enrichment", x=NULL) -> boxes
```

# Visualize similarities between cell responses across conditions
```{r echo=FALSE, fig.width=6, fig.height=6}
library("facetscales")

merged_res_fp <- map(de_combos, function(x) {
  x %>% bind_rows() %>% 
  dplyr::select(gene, cluster_id, logFC, p_val) %>%
  mutate(score=logFC*-log10(p_val))
}) %>% bind_rows(.id="id")

merged_res_fp %>% 
  group_by(gene, cluster_id) %>%
  summarise(
    logFC = median(logFC),
    p_val = median(p_val),
    ) %>% 
  mutate(score=logFC*-log10(p_val))

res_r50$table[[1]] %>% 
  bind_rows() %>% 
  dplyr::select(gene, cluster_id, logFC, p_val) %>%
  mutate(score=logFC*-log10(p_val)) %>%
  left_join(merged_res_fp %>% 
              group_by(gene, cluster_id) %>% 
              summarise(
                logFC = median(logFC),
                p_val = median(p_val)
                ) %>% 
              mutate(score=logFC*-log10(p_val)),
            by = c("gene", "cluster_id")) %>%
  transmute(f_r50 = rank(score.x), s_fgf = score.y, s_r50 = score.x, f_fgf = rank(score.y), gene, cluster_id) %>% 
  split(., .$cluster_id) -> data

map(data, ~ corrr::correlate(.x %>% dplyr::select(contains("f_")), method="spearman")) %>%
  bind_rows(.id = "id") %>% 
  dplyr::select(1,3) %>%
  na.omit() %>% 
  as_tibble() %>%
  rename(f_r50="cor") -> cor

map(data, ~ dplyr::select(.x, c(gene, contains(c("f_","s_"))))) %>%
  bind_rows(.id="id") %>%
  group_by(id) %>%
  mutate(label = case_when(f_r50 < 100 & f_fgf < 100 & !grepl("Rik", gene) ~ gene,
                           f_r50 > max(f_r50)-100 & f_fgf > max(f_fgf)-100 & !grepl("Rik", gene) ~ gene,
                           f_r50 < 100 & f_fgf > max(f_fgf)-100 & !grepl("Rik", gene) ~ gene,
                           f_r50 > max(f_r50)-100 & f_fgf < 100 & !grepl("Rik", gene) ~ gene,
                           T ~ NA_character_)) -> plot

plot %>% 
  #filter(id %in% c("Astro", "Olig", "Tany")) %>%
  ggplot() + 
  aes(x=s_r50, y = s_fgf, label=label) + 
  geom_point(alpha=0.25) +
  geom_hline(yintercept = 0, linetype = "dashed", color="red") +
  geom_vline(xintercept = 0, linetype="dashed", color="red") +
  facet_wrap(. ~ id, ncol = 4, scales = "free") +
  theme_classic() +
  labs(x="R50E v PF", y="FGF v PF") -> genescores

genescores
```

```{r}
plot %>% 
  #filter(id %in% c("Astro", "Olig", "Tany")) %>%
  ggplot() +
  aes(x=f_r50, y = f_fgf, label=label) +
  geom_density_2d_filled(contour_var = "ndensity") +
  facet_wrap(. ~ id, ncol = 4) +
  geom_label(data = cor, alpha=0.5,
             aes(label=paste0("Cor: ", round(cor,3))), 
             x = 40000, y = -Inf, hjust=1, vjust=0,
             inherit.aes = FALSE) +
  labs(fill="Scaled\nDensity") +
  theme_classic() +
  labs(x=NULL, y=NULL) +
  guides(fill = guide_colorsteps()) -> corplots
corplots
```

```{r fig.width=15, fig.height=4}
simil <- cowplot::plot_grid(genescores, heat, boxes,  axis = "tb", align="hv", rel_widths = c(1,1.5,1), nrow=1)
ggsave("diff_r50.png", width = 15, height=4)
```

```{r fig.height=4, fig.width=6}
pdf("genescores.pdf", width = 6,height = 4)
genescores
dev.off()
```
```{r}
pdf("mapk_heat.pdf", width = 6,height = 8)
heat
dev.off()
```
```{r}
pdf("mapk_boxes.pdf", width = 5,height = 8)
boxes
dev.off()
```

```{r fig.width=20, fig.height=5}
left <- cowplot::plot_grid(by_ct, by_trt, diff_ab, de_comp, nrow=1, axis = "tb", align="v", rel_widths = c(1,1,.75), labels = "auto", scale=0.9)
ggsave("top_r50.pdf", w=15, h=5)
```

```{r fig.width=20, fig.height=5}
cowplot::plot_grid(simil, heat, axis="t", align="hv", nrow = 1, scale=c(1,.5))#, labels=c(NA, "d"), scale=c(1,0.95))

```

```{r}
drake::loadd(varplot_wgcna_complete_F_2_15)

purrr::map(varplot_wgcna_complete_F_2_15$mes$Astro, function(x) {
  data.frame(me=as.numeric(x),
             trt = factor(varplot_wgcna_complete_F_2_15$meta$Astro$trt),
             id = factor(varplot_wgcna_complete_F_2_15$meta$Astro$orig.ident)) -> df
  mod <- lme4::lmer(me ~ trt + (1 | id), data = df) 
  pairwise <- emmeans::emmeans(mod, pairwise ~ trt)
  return(pairwise)
}) -> anova_astro

map(anova_astro, ~.x$contrasts %>% 
      data.frame()) %>% 
  bind_rows(.id="id") %>% 
  mutate(padj = p.adjust(`p.value`)) %>%
  write_csv("~/rausch_brown_r50/astro_mods.csv")

map(anova_astro, ~.x$contrasts %>% 
      data.frame()) %>% 
  bind_rows(.id="id") %>% 
  mutate(padj = p.adjust(`p.value`)) %>% 
  arrange(padj) -> anova_astro_df
```

```{r fig.width=12, fig.height=4}
varplot_wgcna_complete_F_2_15$mes$Astro %>% 
  select(ME5, ME11, ME13) %>% 
  mutate(trt = factor(varplot_wgcna_complete_F_2_15$meta$Astro$trt),
         id = factor(varplot_wgcna_complete_F_2_15$meta$Astro$orig.ident)) %>% 
  pivot_longer(c(-trt, -id)) -> df
  
df[sample(nrow(df)),] %>% 
  ggplot() +
  aes(x=trt, y=value, fill = id) +
  ggbeeswarm::geom_quasirandom(alpha=0.5, shape=21) +
  facet_wrap(.~name, scales = "free") +
  labs(x=NULL, y = "Module Expression") +
  theme_bw() +
  NoLegend()
```
# Read in gene sets
```{r}
lps1 <- read_tsv(("/projects/dylan/bentsen-rausch-2019/data/lps1.txt"))
mcao1 <- read_tsv(("/projects/dylan/bentsen-rausch-2019/data/mcao1.txt"))
mcao3 <- read_tsv(("/projects/dylan/bentsen-rausch-2019/data/mcao_d3.txt"))
mcao7 <- read_tsv(("/projects/dylan/bentsen-rausch-2019/data/mcaod7.txt"))
ni <- readxl::read_xlsx(("/projects/dylan/bentsen-rausch-2019/data/neur_astro_induce.xlsx"))
na <- readxl::read_xlsx(("/projects/dylan/bentsen-rausch-2019/data/synaptic_activity_induced.xlsx"))

ni %>%
  dplyr::select(gene_name, `Fold Change`, padj_deseq2) %>%
  filter(`Fold Change` > 2, padj_deseq2 < 0.05) -> ni
na %>%
  dplyr::select(gene_name, Fold_Change, DESeq2_padj) %>%
  filter(Fold_Change > 2, DESeq2_padj < 0.05) -> na


mcao1 %>%
  filter(logFC < (-2)) %>%
  arrange(logFC) %>%
  distinct(Gene.symbol) %>%
  filter(!grepl("///", Gene.symbol)) -> mcao_gene

lps1 %>%
  filter(logFC < (-2)) %>%
  arrange(logFC) %>%
  distinct(Gene.symbol) %>%
  filter(!grepl("///", Gene.symbol)) -> lps_gene

mcao3 %>%
  filter(logFC < (-2)) %>%
  arrange(logFC) %>%
  distinct(Gene.symbol) %>%
  filter(!grepl("///", Gene.symbol)) -> mcao3_gene

mcao7 %>%
  filter(logFC < (-2)) %>%
  arrange(logFC) %>%
  distinct(Gene.symbol) %>%
  filter(!grepl("///", Gene.symbol)) -> mcao7_gene
```

# Filter gene sets
```{r}
intersect(lps_gene$Gene.symbol, mcao_gene$Gene.symbol) -> panreact
lps_uniq <- lps_gene$Gene.symbol[!lps_gene$Gene.symbol %in% mcao_gene$Gene.symbol]
mcao_uniq <- mcao_gene$Gene.symbol[!mcao_gene$Gene.symbol %in% lps_gene$Gene.symbol]
mcao3_uniq <- mcao3_gene$Gene.symbol[!mcao3_gene$Gene.symbol %in% lps_gene$Gene.symbol]
mcao7_uniq <- mcao7_gene$Gene.symbol[!mcao7_gene$Gene.symbol %in% lps_gene$Gene.symbol]
```

# Test module enrichment in gene sets
```{r, fig.height=3, fig.width=5}
varplot_wgcna_complete_F_2_15$gene_mod_df$Astro %>% 
  filter(ME %in% c("ME11","ME13")) %>%
  group_by(ME) %>% 
  mutate(gene = str_to_sentence(gene)) -> astro_mod
astro_mod %>%
  group_split() %>%
  map("gene") -> astro_gene
group_keys(astro_mod) %>% pull(ME) -> mod_names
```

```{r}
lapply(astro_gene, function(x) {
  a <- 1 - phyper(sum(x %in% lps_uniq), length(lps_uniq), 5000, length(x), log.p = F)
  b <- 1 - phyper(sum(x %in% mcao_uniq), length(mcao_uniq), 5000, length(x), log.p = F)
  c <- 1 - phyper(sum(x %in% mcao3_uniq), length(mcao3_uniq), 5000, length(x), log.p = F)
  d <- 1 - phyper(sum(x %in% mcao7_uniq), length(mcao7_uniq), 5000, length(x), log.p = F)
  e <- 1 - phyper(sum(x %in% panreact), length(panreact), 5000, length(x), log.p = F)
  f <- 1 - phyper(sum(x %in% ni$gene_name), length(ni$gene_name), 5000, length(x), log.p = F)
  g <- 1 - phyper(sum(x %in% na$gene_name), length(na$gene_name), 5000, length(x), log.p = F)
  return(data.frame(A1 = a, A2 = b, PAN = e, NeurI = f, `NeurA` = g))
}) %>% bind_rows() -> overlap_test

library(ggpubr)
library(cowplot)
as.data.frame(sapply(overlap_test, function(x) {
  p.adjust(x, n = dim(overlap_test)[2]-1)
  })) -> overlap_test

overlap_test$mod <- mod_names
overlap_pval <- reshape2::melt(overlap_test)
set_plot <- ggplot(overlap_pval, aes(y = mod, 
                                     x = fct_relevel(variable,"NeurA","NeurI","PAN","A2","A1"))) + 
  geom_tile(size = 1, color = "white", fill="grey99") +
  geom_point(aes(size = if_else(-log10(value)<1.3,true = 0, false = -log10(value))), alpha=0.25) + 
  scale_size(name= expression(bold(-log[10] ~ pvalue)), breaks = c(2,4,8)) + 
  scale_fill_manual(values=c("black","red"), guide=F) + coord_flip() + 
  theme_pubr(legend = "right") + xlab(NULL) + ylab(NULL) + 
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
        legend.title = element_text(size=8), 
        legend.text = element_text(size=7, face="bold"),
        axis.text.x = element_text(angle=45, hjust=1))
set_plot
```


```{r}
varplot_wgcna_complete_F_2_15$gene_mod_df$Astro %>% 
  filter(ME=="ME13")
```

```{r}
varplot_wgcna_complete_F_2_15$varpart$Astro %>%
  data.frame() %>%
  rownames_to_column("mod") %>%
  pivot_longer(-mod) %>%
  group_by(name) %>%
  mutate(rank = rank(-value, ties.method = "first")) %>%
  arrange(name, rank) %>%
  ungroup() %>%
  mutate(label = if_else(rank<=3 & name == "trt", true = mod, false=NA_character_),
         name=factor(name)) -> plot_df

ggplot(plot_df) +
  aes(x=fct_relevel(name, "orig.ident","trt","Residuals"), y=value, label=label) +
  geom_violin(aes(fill=name)) + 
  geom_boxplot(width=0.05) +
  geom_text(aes(label=label), na.rm = TRUE, hjust = -0.5, size=3) +
  theme_classic() +
  scale_y_continuous(labels = scales::percent) +
  labs(y="Variance explained", x= NULL) +
  theme(legend.position = "none") -> varplot_astro


data.frame( Expression = varplot_wgcna_complete_F_2_15$mergedMEs$Astro[["newMEs"]][,colnames(varplot_wgcna_complete_F_2_15$mergedMEs$Astro[["newMEs"]])%in%c(plot_df$label[!is.na(plot_df$label)])], 
            trt = varplot_wgcna_complete_F_2_15$meta$Astro[["trt"]], 
            sample = varplot_wgcna_complete_F_2_15$meta$Astro[["orig.ident"]]) %>%
  group_by(trt) %>%
  sample_frac(size = .25) %>% 
  pivot_longer(c(-trt,-sample)) -> GE

ggplot(GE, aes(x=fct_relevel(trt, "PF","R50","FGF"), y=value, fill= sample)) + 
  ggbeeswarm::geom_quasirandom(alpha=0.5, shape=21) + theme_classic() + 
  facet_wrap(.~name) + 
  theme(legend.position = "none") +
  labs(x=NULL, y="ME Value") -> bymod_astro

astro <- cowplot::plot_grid(varplot_astro, bymod_astro, rel_widths = c(1,2))
astro
```

```{r}
varplot_wgcna_complete_F_2_15$varpart$Tany %>%
  data.frame() %>%
  rownames_to_column("mod") %>%
  pivot_longer(-mod) %>%
  group_by(name) %>%
  mutate(rank = rank(-value, ties.method = "first")) %>%
  arrange(name, rank) %>%
  ungroup() %>%
  mutate(label = if_else(rank==1 & name == "trt", true = mod, false=NA_character_),
         name=factor(name)) -> plot_df

ggplot(plot_df) +
  aes(x=fct_relevel(name, "orig.ident","trt","Residuals"), y=value, label=label) +
  geom_violin(aes(fill=name)) + 
  geom_boxplot(width=0.05) +
  geom_text(aes(label=label), na.rm = TRUE, hjust = -0.5, size=3) +
  theme_classic() +
  scale_y_continuous(labels = scales::percent) +
  labs(y="Variance explained", x= NULL) +
  theme(legend.position = "none") -> varplot_tany


data.frame( Expression = varplot_wgcna_complete_F_2_15$mergedMEs$Tany[["newMEs"]][,colnames(varplot_wgcna_complete_F_2_15$mergedMEs$Tany[["newMEs"]])%in%c(plot_df$label[!is.na(plot_df$label)])], 
            trt = varplot_wgcna_complete_F_2_15$meta$Tany[["trt"]], 
            sample = varplot_wgcna_complete_F_2_15$meta$Tany[["orig.ident"]]) %>%
  group_by(trt) %>%
  sample_frac(size = .25) %>% 
  pivot_longer(c(-trt,-sample)) -> GE

ggplot(GE, aes(x=fct_relevel(trt, "PF","R50","FGF"), y=value, fill= sample)) + 
  ggbeeswarm::geom_quasirandom(alpha=0.5, shape=21) + theme_classic() + 
  facet_wrap(.~name) + 
  theme(legend.position = "none") +
  labs(x=NULL, y="ME Value") -> bymod_tany

tany <- cowplot::plot_grid(varplot_tany, bymod_tany, rel_widths = c(1,2))
tany
```

```{r}
cowplot::plot_grid(astro, tany, ncol=1)
```


